---
title: "HW 2: Visualization, Base R"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "Amelia Kim"
  email: "kimam@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`' # no need to change date since this line grabs today's date
output: 
  html_document
---

Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 28 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

**Theme of the assignment**: Let's practice visualization using base R and create cool maps. Load in the `Cars93` dataset by running the following code chunk:

```{r}
library(MASS)
data(Cars93)
```

The library function will load and attach add-on packages that have been previously installed on your machine. This will give you access to the functions and datasets available in the package. In general, you only need to install a package once (although it may require updating from time to time) but you need to access it using the library command every time you start a new session in R (and want the functionality). 

The questions below use the `Cars93` data frame as a source of data. This data frame is part of the MASS package and becomes available after loading it. You can get a view of the top of the data frame by using the head command. For more information, type `?Cars93`. Let's load `Cars93` in,
```{r}
head(Cars93)
```

# Base R
## Task 1
a) Create a Frequency table and Relative Frequency table (latter uses percentages) for the variable `Type`. Make sure to round all digits to 3 digits.
```{r}
#create frequency and relative frequency table
type_frequency <- table(Cars93$Type)
type_relative_frequency <- prop.table(type_frequency) *100

#round to 3 digits
type_frequency <- round(type_frequency, 3)
type_relative_frequency <- round(type_relative_frequency, 3)

#print tables
type_frequency
type_relative_frequency
```
b) Create a frequency bar graph for the variable `Type`. You can have count on the y axis.
```{r}
barplot(type_frequency, main = "Frequency Bar Graph for 'Type'", xlab = "Type", ylab = "Count", col = c("blue", "cyan", "green", "orange", "pink", "red"), cex.lab = 1.2, cex.axis = 1.2, cex.main = 1.2)
```

*	Your axes should be labeled and legible
*	The bars should be of different colors
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

c) Create a pie chart of the variable `Type`.
```{r}
pie(type_frequency, main = "Pie Chart for 'Type'", cex.lab = 1.2, cex.axis = 1.2, cex.main = 1.2)
```

* The plot should include a title using the main argument.
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

d) What is the mode of the variable `Type`? Comment.

#The mode of the variable 'Type' is Midsize.

## Task 2 
a) Create a histogram of the variable `Horsepower`.
```{r}
hist(Cars93$Horsepower, main = "Histogram of Horsepower", xlab = "Horsepower", ylab = "Frequency", col = "lightblue", cex.lab = 1.2, cex.axis = 1.2, cex.main = 1.2)
```

*	Your axes should be labeled
*	The hist body should have a color (your choice, not too dark to obscure lines) 
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes should be set to 1.2

b) Create a second histogram with breaks at 25 HP increments instead of 50.
```{r}
hist(Cars93$Horsepower, main = "Histogram of Horsepower", xlab = "Horsepower", ylab = "Frequency", col = "lightgreen", cex.lab = 1.2, cex.axis = 1.2, cex.main = 1.2, breaks = 25)
```

* Use the breaks argument to define the classes
* Other setting should be as in part (a)

c) Which histogram more closely follows or rule of thumb for the suggested number of classes? Explain.
```{r}
total_data_points <- nrow(Cars93)
suggested_classes <- sqrt(total_data_points)
suggested_classes
```
#The general rule of them for the suggested number of classes is the square root of the total number of scores in the distribution. The output for 'suggested_classes' indicates that around 9.6 classes is suggested, showing that the second histogram with breaks at 25 horsepower more closely follows the rule of thumb compared to the histogram with breaks at 50 horsepower..

## Task 3 
a) What is the type of variable `Passengers`? (Categorical nominal/ordinal or Numerical discrete/continuous?) What happens if we were to make a histogram for `Passengers`? What is the problem with this graph?

#'Passengers' is a numerical discrete variable, consisting of integers to represent how many passengers can fit in car. The problem with making a histogram for 'Passengers' is that the data distribution is skewed. Histograms are usually made with numerical continuous data, where the values don't have gaps in between them.
```{r}
passengers <- c(Cars93$Passengers)
passengers_frequency <- table(passengers)

hist(passengers_frequency, main = "Histogram for 'Passengers'", xlab = "Passengers",
        ylab = "Frequency", cex.lab = 1.2, cex.axis = 1.2, cex.main = 1.2, col = "lightpink")
```

b) Correct the problem in part (a) by creating a more appropriate visualization. Make sure to label your axes, place a title, choose a custom color as in the previous graphs. (Bonus: use factor() with levels option to display x axis values with no observations on the plot.)
```{r}
passengers <- c(Cars93$Passengers)
passengers_frequency <- table(factor(passengers, levels = c("2","3","4","5", "6", "7","8")))

barplot(passengers_frequency, main = "Histogram for 'Passengers'", xlab = "Passengers",
        ylab = "Frequency", cex.lab = 1.2, cex.axis = 1.2, cex.main = 1.2, col = "orange")
```

## Task 4 
a) Make a histogram of the variable `Length` (the default breaks are fine). Other aspects of the plot should be as in part (a) in Task 2.
```{r}
length_cars <- c(Cars93$Length)
length_frequency <- table(length_cars)

hist(length_frequency, main = "Histogram for 'Length'", xlab = "Length", ylab = "Frequency", cex.lab = 1.2, cex.axis = 1.2, cex.main = 1.2,, col = "purple")

```

b) What is the shape of the distribution?

#The shape of the distribution is skewed right.

c) Find the mean and median of the variable `Length.` Do these answers meet your expectations from part (b)?

#The mean is greater than the median of the variable 'Length'. This indicates that the data is skewed right, which agrees with my answer from part (b).
```{r}
length_mean <- mean(Cars93$Length)
length_median <- median(Cars93$Length)

length_mean
length_median
```

You are done! ðŸ Don't forget to commit and push your .Rmd file to your Github repository before the due date.


